### 2025-02-19 Code/R/Research/CP-SSA.Rmd
Add CPD

---
title: "CP-SSA"
output: html_document
---

```{r setup, include=FALSE}
library(Rssa)
library(rTensor)
library(ggplot2)
library(dplyr)
library(tidyr)
# library(parallel)
# library(doSNOW)
library(future.apply)
library(progressr)

setwd(rprojroot::find_rstudio_root_file())
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
source(file = "./Research/TSSA.R")
```

# CP-SSA const test
## Finding best window lengths
### CP-SSA
```{r}
N <- 29
s <- rep(3, N)
groups <- list(1)
sd <- 1
R <- 200

Is <- seq(from = 1, to = N, by = 1)
Ls <- seq(from = 1, to = N, by = 1)

correct_dims <- which(outer(Is, Ls, (\(i, l) ifelse((i >= l) &
                                                      (N - i - l + 2 > 0), TRUE, FALSE
))), arr.ind = TRUE)

len_out <- nrow(correct_dims)

tens_dims <- matrix(ncol = 3, nrow = len_out)
colnames(tens_dims) <- c("I", "L", "J")
tens_dims[, "I"] <- Is[correct_dims[, 1]]
tens_dims[, "L"] <- Ls[correct_dims[, 2]]
tens_dims[, "J"] <- N - tens_dims[, "I"] - tens_dims[, "L"] + 2

tens_dims <- tens_dims[order(tens_dims[, 1], tens_dims[, 2], decreasing = TRUE), ]

findL_cp_ssa <- function() {
  noise <- rnorm(N, sd = sd)
  x <- s + noise

  rec <- apply(tens_dims, 1, function(window)
    tens_ssa_reconstruct(
      x,
      I = window["I"],
      L = window["L"],
      groups = groups,
      decomp = "cp",
      status = FALSE
    ))
  sapply(rec, function(r)
    mean(abs(Reduce("+", r) - s)^2))
}

plan(multisession)
# plan(sequential)
if (!progressr:::register_global_progression_handler("query"))
  progressr:::register_global_progression_handler("add")

const_rmse_L <- function(R) {
  p <- progressor(R)
  future_replicate(R, {
    p()
    findL_cp_ssa()
  }, simplify = "array", future.seed = 5) |> rowMeans() |> sqrt()
}

const_rmse_L_res <- cbind(tens_dims, const_rmse_L(R))
save(const_rmse_L_res, file = "./Research/comp_results/cp_const_L,RData")
const_rmse_L_res[order(const_rmse_L_res[, 4], decreasing = FALSE), ] |> head() |> print()
```

### Basic SSA
```{r}
N <- 29
s <- rep(3, N)
groups <- list(1)
sd <- 1
R <- 200

Ls <- seq(from = 1, to = N, by = 1)

findL_basic_ssa <- function() {
  noise <- rnorm(N, sd = sd)
  x <- s + noise

  rec <- lapply(Ls, function(L)
    reconstruct(ssa(x, L = L), groups = groups))
  sapply(rec, function(r)
    mean(abs(Reduce("+", r) - s)^2))
}

plan(multisession)
# plan(sequential)
if (!progressr:::register_global_progression_handler("query"))
  progressr:::register_global_progression_handler("add")

const_rmse_L_ssa <- function(R) {
  p <- progressor(R)
  future_replicate(R, {
    p()
    findL_basic_ssa()
  }, simplify = "array", future.seed = 5) |> rowMeans() |> sqrt()
}

const_rmse_L_ssa_res <- cbind(Ls, const_rmse_L_ssa(R))
save(const_rmse_L_ssa_res, file = "./Research/comp_results/basic_const_L,RData")
const_rmse_L_ssa_res[order(const_rmse_L_ssa_res[, 2], decreasing = FALSE), ] |> head() |> print()
```

### Comparing CP-SSA and Basic SSA on best windows
```{r}
N <- 29
s <- rep(3, N)
L_ssa <- 12
I_tens <- 18
L_tens <- 12
groups <- list(1)
sd <- 1
R <- 1000

compare_cp_ssa <- function() {
  noise <- rnorm(N, sd = sd)
  x <- s + noise

  rec <- list(
    ssa = reconstruct(ssa(x, L = L_ssa), groups = groups),
    cp_ssa = tens_ssa_reconstruct(
      x,
      I = I_tens,
      L = L_tens,
      groups = groups,
      decomp = "cp",
      status = FALSE
    )
  )
  sapply(rec, function(r)
    mean(abs(Reduce("+", r) - s)^2))
}

plan(multisession)
# plan(sequential)
if (!progressr:::register_global_progression_handler("query"))
  progressr:::register_global_progression_handler("add")

const_rmse <- function(R) {
  p <- progressor(R)
  future_replicate(R, {
    p()
    compare_cp_ssa()
  }, simplify = "array", future.seed = 5) |> rowMeans() |> sqrt()
}

const_rmse_comp_res <- const_rmse(R)
save(L_ssa, I_tens, L_tens, const_rmse_comp_res, file = "./Research/comp_results/cp_const_comp,RData")
print(const_rmse_comp_res)
```