### 2025-02-19 Code/R/Research/Eigenvalues.Rmd

---
title: "Eigenvalues"
output: html_document
---

```{r, setup, include=FALSE}
library(Rssa)
library(rTensor)
library(ggplot2)
library(gridExtra)
library(factoextra)
library(dplyr)
library(tidyr)

knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

# Функции для тензоризации SSA
```{r}
source(file = "./Research/TSSA.R")
```

```{r}
N <- 20; Q <- 4; L <- 10; A <- c(1, 1, 1, 3)
s_const <- matrix(rep(A, N * Q / length(A)), ncol = N) |> t()
tens3(s_const, L = L, kind = "HO-MSSA") |> 
  hosvd_mod(status = FALSE) |>
  get_tensor_eigenvalues() |>
  lapply(zapsmall) |> 
  print()
print(sqrt(mean(A^2)) * sqrt(Q * L * (N - L + 1)))
```