### 2025-02-27 Code/R/Research/CP-SSA.Rmd
Dividing impulse signal

---
title: "CP-SSA"
output: html_document
---

```{r setup, include=FALSE}
library(Rssa)
library(rTensor)
library(ggplot2)
library(dplyr)
library(tidyr)
library(future.apply)
library(progressr)

setwd(rprojroot::find_rstudio_root_file())
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
source(file = "./Research/TSSA.R")
```

```{r enable progressr, eval = FALSE}
# This command should be executed in R console, not as a chunk
# Otherwise it fails
progressr::handlers(global = TRUE)
```

```{r}
gen_nswn <- function(n, lambda = 1, min_sd = 0.1, max_sd = 1.5) {
  step_change_times <- cumsum(rexp(n = floor(n * lambda), rate = lambda))
  while (tail(step_change_times, 1) < n) {
    step_change_times <- c(step_change_times, cumsum(rexp(
      n = floor(n * lambda), rate = lambda
    )))
  }
  step_change_times <- step_change_times[step_change_times < n]

  step_variances <- runif(length(step_change_times) + 1, min = min_sd, max = max_sd)

  variance_modulation <- rep(step_variances[1], n)
  for (i in seq_along(step_change_times)) {
    variance_modulation[seq(from = ceiling(step_change_times[i]),
                            to = n,
                            by = 1)] <- step_variances[i + 1]
  }
  plot(variance_modulation, type = "l")
  rnorm(n) * variance_modulation
}
```

# CP-SSA numerical test functions
## Finding best window lengths
### CP-SSA
```{r}
findL_cp_ssa <- function() {
  noise <- rnorm(N, sd = sd)
  # noise <- gen_nswn(N, lambda = 10 / N)
  x <- s + noise

  rec <- apply(tens_dims, 1, function(L)
    tens_ssa_reconstruct(
      x,
      L[1:2],
      groups = groups,
      decomp = "cp",
      status = FALSE,
      cp_span = cp_span,
      tol = tol,
      max_iter = max_iter
    ))
  sapply(rec, function(r)
    mean(abs(Reduce("+", r) - s)^2))
}
```

### Basic SSA
```{r}
findL_basic_ssa <- function() {
  noise <- rnorm(N, sd = sd)
  # noise <- gen_nswn(N, lambda = 10 / N)
  x <- s + noise

  rec <- lapply(Ls, function(L)
    reconstruct(ssa(x, L = L), groups = groups))
  sapply(rec, function(r)
    mean(abs(Reduce("+", r) - s)^2))
}
```

### MSSA
```{r}
findL_mssa <- function() {
  noise <- rnorm(N, sd = sd)
  # noise <- gen_nswn(N, lambda = 10 / N)
  x <- matrix(s + noise, ncol = amp_period)

  rec <- lapply(Ls, function(L)
    reconstruct(ssa(x, L = L, kind = "mssa"), groups = groups))
  sapply(rec, function(r)
    mean(abs(Reduce("+", r) - s)^2))
}
```

### HO-MSSA
```{r}
findL_homssa <- function() {
  noise <- rnorm(N, sd = sd)
  # noise <- gen_nswn(N, lambda = 10 / N)
  x <- matrix(s + noise, ncol = amp_period)

  rec <- lapply(Ls, function(L)
    tens_mssa_reconstruct(
      x,
      L = L,
      groups = groups,
      groups3 = groups3,
      decomp = "hooi",
      status = FALSE,
      max_iter = max_iter,
      tol = tol
    ))
  sapply(rec, function(r)
    mean(abs(Reduce("+", r) - s)^2))
}
```

## Comparing CP-SSA and Basic SSA on best windows
```{r}
compare_cp_ssa <- function() {
  noise <- rnorm(N, sd = sd)
  # noise <- gen_nswn(N, lambda = 10 / N)
  x <- s + noise

  rec <- list(
    ssa = reconstruct(ssa(x, L = L_ssa), groups = groups_ssa),
    cp_ssa = tens_ssa_reconstruct(
      x,
      L = L_tens,
      groups = groups_tens,
      decomp = "cp",
      status = FALSE,
      cp_span = cp_span,
      tol = tol,
      max_iter = max_iter
    )
  )
  sapply(rec, function(r)
    mean(abs(Reduce("+", r) - s)^2))
}
```

## Comparing Basic SSA, MSSA, CP-SSA, and HO-MSSA on best windows
```{r}
compare_all_ssa <- function() {
  noise <- rnorm(N, sd = sd)
  # noise <- gen_nswn(N, lambda = 10 / N)
  x <- s + noise
  x_mat <- matrix(x, ncol = amp_period)

  rec <- list(
    ssa = reconstruct(ssa(x, L = L_ssa), groups = groups_ssa),
    mssa = reconstruct(ssa(x_mat, L = L_mssa, kind = "mssa"), groups = groups_mssa),
    cp_ssa = tens_ssa_reconstruct(
      x,
      L = L_cp,
      groups = groups_cp,
      decomp = "cp",
      status = FALSE,
      cp_span = cp_span,
      tol = tol,
      max_iter = max_iter
    ),
    ho_mssa = tens_mssa_reconstruct(
      x_mat,
      L = L_homssa,
      groups = groups_homssa,
      groups3 = groups3,
      decomp = "hooi",
      status = FALSE,
      tol = tol,
      max_iter = max_iter
    )
  )
  sapply(rec, function(r)
    mean(abs(Reduce("+", r) - s)^2))
}
```

# Extraction precision 
## Const ts
### Best $(l, J)$ for CP-SSA 
```{r, warning=FALSE}
N <- 28
s <- rep(3, N)
groups <- list(1)
sd <- 1
R <- 200
cp_span <- "mean"
# cp_span <- "last"
tol <- 1e-7
max_iter = 25

ls <- seq(from = 1,
          to = (N + 1) %/% 2,
          by = 1)
Js <- seq(from = 1, to = N, by = 1)

correct_dims <- which(outer(ls, Js, (\(l, j) ifelse((l >= j) &
                                                      (l - j + 1 >= 1), TRUE, FALSE
))), arr.ind = TRUE)

len_out <- nrow(correct_dims)

tens_dims <- matrix(ncol = 3, nrow = len_out)
colnames(tens_dims) <- c("l", "J", "K")
tens_dims[, "l"] <- ls[correct_dims[, 1]]
tens_dims[, "J"] <- Js[correct_dims[, 2]]
tens_dims[, "K"] <- tens_dims[, "l"] - tens_dims[, "J"] + 1

tens_dims <- tens_dims[order(tens_dims[, 1], tens_dims[, 2], decreasing = TRUE), ]

plan(multisession)
# plan(sequential)

const_rmse_L <- function(R) {
  p <- progressor(R)
  future_replicate(R, {
    p()
    findL_cp_ssa()
  }, simplify = "array", future.seed = 5) |> rowMeans() |> sqrt()
}

const_rmse_L_res <- cbind(tens_dims, const_rmse_L(R))
save(const_rmse_L_res, file = "./Research/comp_results/cp_const_L.RData")
const_rmse_L_res[order(const_rmse_L_res[, 4], decreasing = FALSE), ] |> head() |> print()
# last: 14  9 6 0.3580338
# mean: 12  8  5 0.3534230
```

### Best L for basic SSA
```{r}
N <- 28
s <- rep(3, N)
groups <- list(1)
sd <- 1
R <- 200

Ls <- seq(from = 1, to = (N + 1) %/% 2, by = 1)

plan(multisession)
# plan(sequential)

const_rmse_L_ssa <- function(R) {
  p <- progressor(R)
  future_replicate(R, {
    p()
    findL_basic_ssa()
  }, simplify = "array", future.seed = 5) |> rowMeans() |> sqrt()
}

const_rmse_L_ssa_res <- cbind(Ls, const_rmse_L_ssa(R))
save(const_rmse_L_ssa_res, file = "./Research/comp_results/basic_const_L.RData")
const_rmse_L_ssa_res[order(const_rmse_L_ssa_res[, 2], decreasing = FALSE), ] |> head() |> print()
# 12 0.3026873
```

### Comparing CP-SSA and Basic SSA on best windows
```{r}
N <- 28
s <- rep(3, N)
L_ssa <- 12
# L_tens <- c(14, 9)
L_tens <- c(12, 8)
groups_ssa <- list(1)
groups_tens <- list(1)
sd <- 1
R <- 1000
tol <- 1e-7
max_iter = 25
cp_span = "mean"

plan(multisession)
# plan(sequential)

const_rmse <- function(R) {
  p <- progressor(R)
  future_replicate(R, {
    p()
    compare_cp_ssa()
  }, simplify = "array", future.seed = 5) |> rowMeans() |> sqrt()
}

const_rmse_comp_res <- const_rmse(R)
save(L_ssa, L_tens, const_rmse_comp_res, file = "./Research/comp_results/cp_const_comp.RData")
print(const_rmse_comp_res)
```

## Cos ts
### Best $(I, L)$ for CP-SSA 
```{r}
N <- 28
freq <- 1 / 4
s <- 2 * cos(2 * pi * (1:N) * freq)
groups <- list(1)
max.r <- max(sapply(groups, max))
sd <- 1
R <- 200
cp_span <- "last"
tol <- 1e-7
max_iter = 25

ls <- seq(from = max.r,
          to = (N + 1) %/% 2,
          by = 1)
Js <- seq(from = max.r, to = max(ls) - max.r + 1, by = 1)

correct_dims <- which(outer(ls, Js, (\(l, j) ifelse((l >= j) &
                                                      (l - j + 1 >= max.r), TRUE, FALSE
))), arr.ind = TRUE)

len_out <- nrow(correct_dims)

tens_dims <- matrix(ncol = 3, nrow = len_out)
colnames(tens_dims) <- c("l", "J", "K")
tens_dims[, "l"] <- ls[correct_dims[, 1]]
tens_dims[, "J"] <- Js[correct_dims[, 2]]
tens_dims[, "K"] <- tens_dims[, "l"] - tens_dims[, "J"] + 1

tens_dims <- tens_dims[order(tens_dims[, 1], tens_dims[, 2], decreasing = TRUE), ]

plan(multisession)
# plan(sequential)

cos_rmse_L <- function(R) {
  p <- progressor(R)
  future_replicate(R, {
    p()
    findL_cp_ssa()
  }, simplify = "array", future.seed = 5) |> rowMeans() |> sqrt()
}

cos_rmse_L_res <- cbind(tens_dims, cos_rmse_L(R))
# save(cos_rmse_L_res, file = "./Research/comp_results/cp_cos_L.RData")
cos_rmse_L_res[order(cos_rmse_L_res[, 4], decreasing = FALSE), ] |> head() |> print()
# groups = list(1):   14 12  3 0.6182172
# groups = list(1:2): 14  7  8 0.5012795
# groups = list(1:3): 14  9  6 0.6573032
```

### Best L for basic SSA
```{r}
N <- 28
freq <- 1 / 4
s <- 2 * cos(2 * pi * (1:N) * freq)
groups <- list(1:2)
max.r <- max(sapply(groups, max))
sd <- 1
R <- 200

Ls <- seq(from = max.r, to = (N + 1) %/% 2, by = 1)

plan(multisession)
# plan(sequential)

cos_rmse_L_ssa <- function(R) {
  p <- progressor(R)
  future_replicate(R, {
    p()
    findL_basic_ssa()
  }, simplify = "array", future.seed = 5) |> rowMeans() |> sqrt()
}

cos_rmse_L_ssa_res <- cbind(Ls, cos_rmse_L_ssa(R))
save(cos_rmse_L_ssa_res, file = "./Research/comp_results/basic_cos_L.RData")
cos_rmse_L_ssa_res[order(cos_rmse_L_ssa_res[, 2], decreasing = FALSE), ] |> head() |> print()
# 11 0.4400179
```

### Comparing CP-SSA and Basic SSA on best windows
```{r}
N <- 28
freq <- 1 / 4
s <- 2 * cos(2 * pi * (1:N) * freq)
groups_ssa <- list(1:2)
groups_tens <- list(1:2)
L_ssa <- 11
L_tens <- c(14, 12)
sd <- 1
R <- 1000
tol <- 1e-7
max_iter = 25

plan(multisession)
# plan(sequential)

cos_rmse <- function(R) {
  p <- progressor(R)
  future_replicate(R, {
    p()
    compare_cp_ssa()
  }, simplify = "array", future.seed = 5) |> rowMeans() |> sqrt()
}

cos_rmse_comp_res <- cos_rmse(R)
save(L_ssa, L_tens, cos_rmse_comp_res, file = "./Research/comp_results/cp_cos_comp.RData")
print(cos_rmse_comp_res)
```

## Example from paper
### Best $(l, J)$ for CP-SSA 
```{r, warning=FALSE}
N <- 6000
times <- seq(from = 0,
             to = 1,
             length.out = N)
s <- 2 * exp(-0.2 * pi * 50 * times) * sin(2 * pi * 50 * times) + sin(2 * pi * 10 * times)

groups <- list(1:4)
max.r <- max(sapply(groups, max))
sd <- 0.5
R <- 200
# cp_span <- "mean"
cp_span <- "last"
tol <- 1e-7
max_iter = 25

ls <- seq(from = 2520,
          to = (N + 1) %/% 2,
          by = 120)
Js <- seq(from = 840, to = max(ls) - max.r + 1, by = 120)

correct_dims <- which(outer(ls, Js, (\(l, j) ifelse((l >= j) &
                                                      (l - j + 1 <= j) &
                                                      (l - j + 1 >= max.r), TRUE, FALSE
))), arr.ind = TRUE)

len_out <- nrow(correct_dims)

tens_dims <- matrix(ncol = 3, nrow = len_out)
colnames(tens_dims) <- c("l", "J", "K")
tens_dims[, "l"] <- ls[correct_dims[, 1]]
tens_dims[, "J"] <- Js[correct_dims[, 2]]
tens_dims[, "K"] <- tens_dims[, "l"] - tens_dims[, "J"] + 1

tens_dims <- tens_dims[order(tens_dims[, 1], tens_dims[, 2], decreasing = TRUE), ]

plan(multisession)
# plan(sequential)

expmod_rmse_L <- function(R) {
  p <- progressor(R)
  future_replicate(R, {
    p()
    findL_cp_ssa()
  }, simplify = "array", future.seed = 5) |> rowMeans() |> sqrt()
}

expmod_rmse_L_res <- cbind(tens_dims, expmod_rmse_L(R))
save(expmod_rmse_L_res, file = "./Research/comp_results/cp_expmod_L.RData")
expmod_rmse_L_res[order(expmod_rmse_L_res[, 4], decreasing = FALSE), ] |> head() |> print()
# 3000 2760 241 0.05714844
```

```{r, warning=FALSE}
N <- 6000
times <- seq(from = 0,
             to = 1,
             length.out = N)
s <- 2 * exp(-0.2 * pi * 50 * times[1:(N / 5)]) * sin(2 * pi * 50 * times) + sin(2 * pi * 10 * times)

groups <- list(1:4)
max.r <- max(sapply(groups, max))
sd <- 0.5
R <- 200
# cp_span <- "mean"
cp_span <- "last"
tol <- 1e-7
max_iter = 15

ls <- seq(from = 2520,
          to = (N + 1) %/% 2,
          by = 120)
Js <- seq(from = 840, to = max(ls) - max.r + 1, by = 120)

correct_dims <- which(outer(ls, Js, (\(l, j) ifelse((l >= j) &
                                                      (l - j + 1 <= j) &
                                                      (l - j + 1 >= max.r), TRUE, FALSE
))), arr.ind = TRUE)

len_out <- nrow(correct_dims)

tens_dims <- matrix(ncol = 3, nrow = len_out)
colnames(tens_dims) <- c("l", "J", "K")
tens_dims[, "l"] <- ls[correct_dims[, 1]]
tens_dims[, "J"] <- Js[correct_dims[, 2]]
tens_dims[, "K"] <- tens_dims[, "l"] - tens_dims[, "J"] + 1

tens_dims <- tens_dims[order(tens_dims[, 1], tens_dims[, 2], decreasing = TRUE), ]

plan(multisession)
# plan(sequential)

pulse_rmse_L <- function(R) {
  p <- progressor(R)
  future_replicate(R, {
    p()
    findL_cp_ssa()
  }, simplify = "array", future.seed = 5) |> rowMeans() |> sqrt()
}

pulse_rmse_L_res <- cbind(tens_dims, pulse_rmse_L(R))
save(pulse_rmse_L_res, file = "./Research/comp_results/cp_pulse_L.RData")
pulse_rmse_L_res[order(pulse_rmse_L_res[, 4], decreasing = FALSE), ] |> head() |> print()
```

### Best L for SSA
```{r}
N <- 6000
times <- seq(from = 0,
             to = 1,
             length.out = N)
s <- 2 * exp(-0.2 * pi * 50 * times) * sin(2 * pi * 50 * times) + sin(2 * pi * 10 * times)

Ls <- seq(from = 1920, to = N %/% 2, by = 120)
groups <- list(1:2, 3:4)
sd <- 0.5
R <- 200

plan(multisession)
# plan(sequential)

expmod_rmse_L_ssa <- function(R) {
  p <- progressor(R)
  future_replicate(R, {
    p()
    findL_basic_ssa()
  }, simplify = "array", future.seed = 5) |> rowMeans() |> sqrt()
}

expmod_rmse_L_ssa_res <- cbind(Ls, pulse_rmse_L_ssa(R))
save(expmod_rmse_L_ssa_res, file = "./Research/comp_results/basic_expmod_L.RData")
expmod_rmse_L_ssa_res[order(expmod_rmse_L_ssa_res[, 2], decreasing = FALSE), ] |> head() |> print()
# 2280 0.02389376
```

```{r}
N <- 6000
times <- seq(from = 0,
             to = 1,
             length.out = N)
s <- 2 * exp(-0.2 * pi * 50 * times[1:(N / 5)]) * sin(2 * pi * 50 * times) + sin(2 * pi * 10 * times)

Ls <- seq(from = 1920, to = N %/% 2, by = 120)
groups <- list(1:2, 3:4)
sd <- 0.5
R <- 200

plan(multisession)
# plan(sequential)

pulse_rmse_L_ssa <- function(R) {
  p <- progressor(R)
  future_replicate(R, {
    p()
    findL_basic_ssa()
  }, simplify = "array", future.seed = 5) |> rowMeans() |> sqrt()
}

pulse_rmse_L_ssa_res <- cbind(Ls, pulse_rmse_L_ssa(R))
save(pulse_rmse_L_ssa_res, file = "./Research/comp_results/basic_pulse_L.RData")
pulse_rmse_L_ssa_res[order(pulse_rmse_L_ssa_res[, 2], decreasing = FALSE), ] |> head() |> print()
```

### Comparing CP-SSA and Basic SSA
```{r}
N <- 6000
times <- seq(from = 0,
             to = 1,
             length.out = N)
s <- 2 * exp(-0.2 * pi * 50 * times) * sin(2 * pi * 50 * times) + sin(2 * pi * 10 * times)
L_ssa <- 2280
L_tens <- c(3000, 2760)
groups_ssa <- list(1:2, 3:4)
groups_tens <- list(c(1, 3), c(2, 4))
sd <- 0.5
R <- 1000
tol <- 1e-7
max_iter = 25
cp_span = "last"

plan(multisession)
# plan(sequential)

expmod_rmse <- function(R) {
  p <- progressor(R)
  future_replicate(R, {
    p()
    compare_cp_ssa()
  }, simplify = "array", future.seed = 5) |> rowMeans() |> sqrt()
}

expmod_rmse_comp_res <- expmod_rmse(R)
save(L_ssa, L_tens, expmod_rmse_comp_res, file = "./Research/comp_results/cp_expmod_comp.RData")
print(expmod_rmse_comp_res)
```

```{r}
N <- 6000
times <- seq(from = 0,
             to = 1,
             length.out = N)
s <- 2 * exp(-0.2 * pi * 50 * times[1:(N / 5)]) * sin(2 * pi * 50 * times) + sin(2 * pi * 10 * times)
L_ssa <- 1920
L_tens <- c(600, 300)
groups_ssa <- list(1:32)
groups_tens <- list(c(1, 3), c(2, 4))
sd <- 0.5
R <- 1000
tol <- 1e-7
max_iter = 25
cp_span = "last"

plan(multisession)
# plan(sequential)

pulse_rmse <- function(R) {
  p <- progressor(R)
  future_replicate(R, {
    p()
    compare_cp_ssa()
  }, simplify = "array", future.seed = 5) |> rowMeans() |> sqrt()
}

pulse_rmse_comp_res <- pulse_rmse(R)
save(L_ssa, L_tens, pulse_rmse_comp_res,  groups_ssa, file = "./Research/comp_results/cp_pulse_comp.RData")
print(pulse_rmse_comp_res)
```

# Dividing signal from paper and using MSSA
## Finding best windows
### MSSA
```{r}
N <- 6000
times <- seq(from = 0,
             to = 1,
             length.out = N)
freq <- 50
freq2 <- 10
amp_period <- 5
s <- 2 * exp(-0.2 * pi * freq * times[1:(N %/% amp_period)]) * sin(2 * pi * freq * times) +
  sin(2 * pi * freq2 * times)

groups <- list(1:2, 3:4)
max.r <- max(sapply(groups, max)) 
Ls <- seq(from = 120, to = N %/% amp_period - max.r, by = 120)

sd <- 0.5
R <- 200

plan(multisession)
# plan(sequential)

pulse_rmse_L_mssa <- function(R) {
  p <- progressor(R)
  future_replicate(R, {
    p()
    findL_mssa()
  }, simplify = "array", future.seed = 5) |> rowMeans() |> sqrt()
}

pulse_rmse_L_mssa_res <- cbind(Ls, pulse_rmse_L_mssa(R))
save(pulse_rmse_L_mssa_res, file = "./Research/comp_results/mssa_pulse_L.RData")
pulse_rmse_L_mssa_res[order(pulse_rmse_L_mssa_res[, 2], decreasing = FALSE), ] |> head() |> print()
# 840 0.03829296
```

### HO-MSSA
```{r}
N <- 6000
times <- seq(from = 0,
             to = 1,
             length.out = N)
freq <- 50
freq2 <- 10
amp_period <- 5
s <- 2 * exp(-0.2 * pi * freq * times[1:(N %/% amp_period)]) * sin(2 * pi * freq * times) +
  sin(2 * pi * freq2 * times)

Ls <- seq(from = 120, to = (N %/% amp_period) %/% 2, by = 120)
groups <- list(1:2, 3:4)
groups3 <- list(1:3, 1:3)
sd <- 0.5
R <- 200

plan(multisession)
# plan(sequential)

pulse_rmse_L_homssa <- function(R) {
  p <- progressor(R)
  future_replicate(R, {
    p()
    findL_homssa()
  }, simplify = "array", future.seed = 5) |> rowMeans() |> sqrt()
}

pulse_rmse_L_homssa_res <- cbind(Ls, pulse_rmse_L_homssa(R))
save(pulse_rmse_L_homssa_res, file = "./Research/comp_results/homssa_pulse_L.RData")
pulse_rmse_L_homssa_res[order(pulse_rmse_L_homssa_res[, 2], decreasing = FALSE), ] |> head() |> print()
# 360 0.03752320
```