### 2025-03-12 Code/R/Research/CP_MSSA.Rmd
CP noise mixing

---
title: "CP_MSSA"
output: html_document
---

```{r setup, include=FALSE}
library(Rssa)
library(rTensor)
library(ggplot2)
library(dplyr)
library(tidyr)
library(future.apply)
library(progressr)

setwd(rprojroot::find_rstudio_root_file())
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
source(file = "./Research/TSSA.R")
```

```{r enable progressr, eval = FALSE}
# This commands should be executed in R console, not as a chunk
# Otherwise they fail
progressr::handlers(global = TRUE)
progressr::handlers("progress")
```

```{r}
gen_nswn <- function(n, lambda = 1, min_sd = 0.1, max_sd = 1.5) {
  step_change_times <- cumsum(rexp(n = floor(n * lambda), rate = lambda))
  while (tail(step_change_times, 1) < n) {
    step_change_times <- c(step_change_times, cumsum(rexp(
      n = floor(n * lambda), rate = lambda
    )))
  }
  step_change_times <- step_change_times[step_change_times < n]

  step_variances <- runif(length(step_change_times) + 1, min = min_sd, max = max_sd)

  variance_modulation <- rep(step_variances[1], n)
  for (i in seq_along(step_change_times)) {
    variance_modulation[seq(from = ceiling(step_change_times[i]),
                            to = n,
                            by = 1)] <- step_variances[i + 1]
  }
  plot(variance_modulation, type = "l")
  rnorm(n) * variance_modulation
}
```

# CP-SSA numerical test functions
## Finding best window lengths
### CP-MSSA
```{r}
findL_cp_mssa <- function() {
  noise <- rnorm(N * Q, sd = sd)
  # noise <- gen_nswn(N, lambda = 10 / N)
  x <- s + noise

  rec <- lapply(Ls_cp, function(L)
    tens_mssa_reconstruct(
      x,
      L,
      groups = groups_cp,
      decomp = "cp",
      status = FALSE,
      tol = tol,
      max_iter = max_iter,
      start = "svd"
    ))
  sapply(rec, function(r)
    colMeans(abs(Reduce("+", r) - s)^2))
}
```

### MSSA
```{r}
findL_mssa <- function() {
  noise <- rnorm(N * Q, sd = sd)
  # noise <- gen_nswn(N, lambda = 10 / N)
  x <- s + noise

  rec <- lapply(Ls_mssa, function(L)
    reconstruct(ssa(x, L = L, kind = "mssa"), groups = groups_mssa))
  sapply(rec, function(r)
    colMeans(abs(Reduce("+", r) - s)^2))
}
```

### HO-MSSA
```{r}
findL_homssa <- function() {
  noise <- rnorm(N * Q, sd = sd)
  # noise <- gen_nswn(N, lambda = 10 / N)
  x <- s + noise

  rec <- lapply(Ls_ho, function(L)
    tens_mssa_reconstruct(
      x,
      L = L,
      groups = groups_mssa,
      groups3 = groups3,
      decomp = "hooi",
      status = FALSE,
      max_iter = max_iter,
      tol = tol
    ))
  sapply(rec, function(r)
    colMeans(abs(Reduce("+", r) - s)^2))
}
```

## Comparing MV methods on best windows
```{r}
compare_all_mssa <- function() {
  noise <- rnorm(N * Q, sd = sd)
  # noise <- gen_nswn(N, lambda = 10 / N)
  x <- s + noise

  rec <- list(
    mssa = reconstruct(ssa(x, L = L_mssa, kind = "mssa"), groups = groups_mssa),
    cp_mssa = tens_mssa_reconstruct(
      x,
      L_cp,
      groups = groups_cp,
      decomp = "cp",
      status = FALSE,
      tol = tol,
      max_iter = max_iter,
      start = "svd"
    ),
    ho_mssa = tens_mssa_reconstruct(
      x,
      L = L_ho,
      groups = groups_mssa,
      groups3 = groups3,
      decomp = "hooi",
      status = FALSE,
      tol = tol,
      max_iter = max_iter
    )
  )
  sapply(rec, function(r)
    colMeans(abs(Reduce("+", r) - s)^2))
}
```

# Using CP for MSSA
## Setup
```{r}
N <- 71
ones <- rep(1, N)
## 1
# A <- ones %o% seq(25, 35, 2.5)
# w <- rep(1 / 12, 5)
# phi <- ones %o% rep(0, 5)
# save_file_name <- "same_freq"
# groups_mssa <- list(1:2)
# groups3 <- list(1)
# groups_cp <- list(1:2)

## 2
# A <- ones %o% c(30, 20, 25, 32)
# w <- c(1 / 12, 1 / 8, 1/12, 1/8)
# phi <- ones %o% c(0, pi / 2, 2 * pi / 2, 3 * pi / 2)
# save_file_name <- "diff_freq_diff_phase"
# groups_mssa <- list(1:4)
# groups3 <- list(1:2)
# groups_cp <- list(1:8)

## 3
# A <- ones %o% rep(c(30, 28, 26, 24, 22), 2)[1:9]
# w <- rep(1 / 12, 9)
# phi <- ones %o% ((0:8) * 2 * pi / 3)
# save_file_name <- "same_freq_diff_phase"
# groups_mssa <- list(1:2)
# groups3 <- list(1:2)
# groups_cp <- list(1:4)

## 4
# A <- 20 * exp(ones %o% c(-rep(0.05, 9)) * 1:N)
# w <- rep(1 / 12, 9)
# phi <- ones %o% rep(0, 9)
# save_file_name <- "same_exp_same_freq"
# groups_mssa <- list(1:2)
# groups3 <- list(1)
# groups_cp <- list(1:2)

## 5
A <- 20 * exp(ones %o% c(rep(0.05, 7), 0.03, 0.03) * 1:N)
w <- rep(1 / 12, 9)
phi <- ones %o% rep(0, 9)
save_file_name <- "diff_exp_same_freq"
groups_mssa <- list(1:4)
groups3 <- list(1:2)
groups_cp <- list(1:4)

s <- A * cos(2 * pi * 1:N %o% w + phi)
Q <- ncol(s)

max.r <- max(sapply(groups_mssa, max))
max.rcp <- max(sapply(groups_cp, max))

Ls_mssa <- seq(from = max.r,
               to = N - max.r + 1,
               by = 2)
Ls_cp <- seq(from = max.rcp,
             to = N - max.rcp + 1,
             by = 2)
# Ls_cp <- c(20, 40)
Ls_ho <- seq(from = max.r,
                 to = (N + 1) %/% 2,
                 by = 2)

sd <- 5
max_iter <- 25
tol <- 1e-7
```

## Best L for CP-MSSA
```{r}
R <- 200

plan(multisession)
# plan(sequential)

rmse_L_cpmssa <- function(R) {
  p <- progressor(R)
  future_replicate(R, {
    p()
    findL_cp_mssa()
  }, simplify = "array", future.seed = 5) |> apply(1:2, mean) |> sqrt()
}

rmse_L_cpmssa_res <- t(rbind(Ls_cp, rmse_L_cpmssa(R)))
# save(
#   rmse_L_cpmssa_res,
#   file = paste0(
#     "./Research/comp_results/cpmssa_",
#     save_file_name,
#     "_L.RData"
#   )
# )
rmse_L_cpmssa_res[order(rowMeans(rmse_L_cpmssa_res[, -1]), decreasing = FALSE), ] |>
  head() |>
  print()
# 62 1.139832 1.163155 1.211639 1.165164 1.210579 1.182345 1.168282 2.261075 2.241531
# 56 1.102377 1.134577 1.175887 1.130599 1.181898 1.151490 1.109888 2.002356 2.020384
```

## Best L for HO-MSSA
```{r}
R <- 200

plan(multisession)
# plan(sequential)

rmse_L_homssa <- function(R) {
  p <- progressor(R)
  future_replicate(R, {
    p()
    findL_homssa()
  }, simplify = "array", future.seed = 5) |> apply(1:2, mean) |> sqrt()
}

rmse_L_homssa_res <- t(rbind(Ls_ho, rmse_L_homssa(R)))
save(
  rmse_L_homssa_res,
  file = paste0(
    "./Research/comp_results/homssa_",
    save_file_name,
    "_L.RData"
  )
)
rmse_L_homssa_res[order(rowMeans(rmse_L_homssa_res[, -1]), decreasing = FALSE), ] |>
  head() |>
  print()
```

## Best L for MSSA
```{r}
R <- 200

plan(multisession)
# plan(sequential)

rmse_L_mssa <- function(R) {
  p <- progressor(R)
  future_replicate(R, {
    p()
    findL_mssa()
  }, simplify = "array", future.seed = 5) |> apply(1:2, mean) |> sqrt()
}

rmse_L_mssa_res <- t(rbind(Ls_mssa, rmse_L_mssa(R)))
save(
  rmse_L_mssa_res,
  file = paste0("./Research/comp_results/mssa_", save_file_name, "_L.RData")
)
rmse_L_mssa_res[order(rowMeans(rmse_L_mssa_res[, -1]), decreasing = FALSE), ] |>
  head() |>
  print()
```

## Comparing MV methods
```{r}
R <- 1000

L_cp <- rmse_L_cpmssa_res[which.min(rowMeans(rmse_L_cpmssa_res[, -1])), 1]
L_ho <- rmse_L_homssa_res[which.min(rowMeans(rmse_L_homssa_res[, -1])), 1]
L_mssa <- rmse_L_mssa_res[which.min(rowMeans(rmse_L_mssa_res[, -1])), 1]

all_mv_rmse <- function(R) {
  p <- progressor(R)
  future_replicate(R, {
    p()
    compare_all_mssa()
  }, simplify = "array", future.seed = 5) |> apply(1:2, mean) |> sqrt()
}

all_mv_rmse_res <- all_mv_rmse(R)
save(
  L_mssa,
  L_cp,
  L_ho,
  groups_mssa,
  groups_cp,
  groups3,
  all_mv_rmse_res,
  sd,
  file = paste0(
    "./Research/comp_results/all_",
    save_file_name,
    "_mv_comp.RData"
  )
)
print(all_mv_rmse_res)
```