### 2025-03-19 Code/R/Research/TSVD_MSSA.Rmd
tsvd-mssa research

---
title: "TSVD_MSSA"
output: html_document
---

```{r setup, include=FALSE}
library(Rssa)
library(rTensor)
library(ggplot2)
library(dplyr)
library(tidyr)
library(future.apply)
library(progressr)

setwd(rprojroot::find_rstudio_root_file())
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
source(file = "./Research/TSSA.R")
```

```{r enable progressr, eval = FALSE}
# This commands should be executed in R console, not as a chunk
# Otherwise they fail
progressr::handlers(global = TRUE)
progressr::handlers("progress")
```

# TSVD-SSA numerical test functions
## Finding best window lengths
### TSVD-MSSA
```{r}
findL_tsvd_mssa <- function() {
  noise <- rnorm(N * Q, sd = sd)
  x <- s + noise

  rec <- lapply(Ls_tsvd, function(L)
    tens_mssa_reconstruct(
      x,
      L,
      groups = groups_tsvd,
      decomp = "tsvd",
      status = FALSE
    ))
  sapply(rec, function(r)
    colMeans(abs(Reduce("+", r) - s)^2))
}
```

### MSSA
```{r}
findL_mssa <- function() {
  noise <- rnorm(N * Q, sd = sd)
  x <- s + noise

  rec <- lapply(Ls_mssa, function(L)
    reconstruct(ssa(x, L = L, kind = "mssa"), groups = groups_mssa))
  sapply(rec, function(r)
    colMeans(abs(Reduce("+", r) - s)^2))
}
```

### HO-MSSA
```{r}
findL_homssa <- function() {
  noise <- rnorm(N * Q, sd = sd)
  x <- s + noise

  rec <- lapply(Ls_ho, function(L)
    tens_mssa_reconstruct(
      x,
      L = L,
      groups = groups_mssa,
      groups3 = groups3,
      decomp = "hooi",
      status = FALSE,
      max_iter = max_iter,
      tol = tol
    ))
  sapply(rec, function(r)
    colMeans(abs(Reduce("+", r) - s)^2))
}
```

## Comparing MV methods on best windows
```{r}
compare_all_mssa <- function() {
  noise <- rnorm(N * Q, sd = sd)
  x <- s + noise

  rec <- list(
    mssa = reconstruct(ssa(x, L = L_mssa, kind = "mssa"), groups = groups_mssa),
    tsvd_mssa = tens_mssa_reconstruct(
      x,
      L_tsvd,
      groups = groups_tsvd,
      decomp = "tsvd",
      status = FALSE
    ),
    ho_mssa = tens_mssa_reconstruct(
      x,
      L = L_ho,
      groups = groups_mssa,
      groups3 = groups3,
      decomp = "hooi",
      status = FALSE,
      tol = tol,
      max_iter = max_iter
    )
  )
  sapply(rec, function(r)
    colMeans(abs(Reduce("+", r) - s)^2))
}
```

# Using TSVD for MSSA
## Setup
```{r}
N <- 71
ones <- rep(1, N)
## 1
# A <- ones %o% seq(25, 35, 2.5)
# w <- rep(1 / 12, 5)
# phi <- ones %o% rep(0, 5)
# save_file_name <- "same_freq"
# groups_mssa <- list(1:2)
# groups3 <- list(1)
# groups_tsvd <- list(1:2)

## 2
A <- ones %o% c(30, 20, 25, 32)
w <- c(1 / 12, 1 / 8, 1/12, 1/8)
phi <- ones %o% c(0, pi / 2, 2 * pi / 2, 3 * pi / 2)
save_file_name <- "diff_freq_diff_phase"
groups_mssa <- list(1:4)
groups3 <- list(1:2)
groups_tsvd <- list(1:4)

## 3
# A <- ones %o% rep(c(30, 28, 26, 24, 22), 2)[1:9]
# w <- rep(1 / 12, 9)
# phi <- ones %o% ((0:8) * 2 * pi / 3)
# save_file_name <- "same_freq_diff_phase"
# groups_mssa <- list(1:2)
# groups3 <- list(1:2)
# groups_tsvd <- list(1:4)

## 4
# A <- 20 * exp(ones %o% c(-rep(0.05, 9)) * 1:N)
# w <- rep(1 / 12, 9)
# phi <- ones %o% rep(0, 9)
# save_file_name <- "same_exp_same_freq"
# groups_mssa <- list(1:2)
# groups3 <- list(1)
# groups_tsvd <- list(1:2)

# 5
# A <- 20 * exp(ones %o% c(rep(0.05, 7), 0.03, 0.03) * 1:N)
# w <- rep(1 / 12, 9)
# phi <- ones %o% rep(0, 9)
# save_file_name <- "diff_exp_same_freq"
# groups_mssa <- list(1:4)
# groups3 <- list(1:2)
# groups_tsvd <- list(1:4)

## 6
# A <- 25 * exp(ones %o% c(-rep(0.05, 7), -0.03, -0.03) * 1:N)
# w <- rep(1 / 12, 9)
# phi <- ones %o% rep(0, 9)
# save_file_name <- "noise_mixing"
# groups_mssa <- list(1:2)
# groups3 <- list(1)
# groups_tsvd <- list(1:2)

s <- A * cos(2 * pi * 1:N %o% w + phi)
Q <- ncol(s)

max.r <- max(sapply(groups_mssa, max))
max.rtsvd <- max(sapply(groups_tsvd, max))

Ls_mssa <- seq(from = max.r,
               to = N - max.r + 1,
               by = 2)
Ls_tsvd <- seq(from = max.rtsvd,
             to = N - max.rtsvd + 1,
             by = 2)
Ls_ho <- seq(from = max.r,
                 to = (N + 1) %/% 2,
                 by = 2)

sd <- 5
max_iter <- 25
tol <- 1e-7
```

## Best L for TSVD-MSSA
```{r}
R <- 200

plan(multisession)
# plan(sequential)

rmse_L_tsvdmssa <- function(R) {
  p <- progressor(R)
  future_replicate(R, {
    p()
    findL_tsvd_mssa()
  }, simplify = "array", future.seed = 5) |> apply(1:2, mean) |> sqrt()
}

rmse_L_tsvdmssa_res <- t(rbind(Ls_tsvd, rmse_L_tsvdmssa(R)))
save(
  rmse_L_tsvdmssa,
  file = paste0(
    "./Research/comp_results/tsvdmssa_",
    save_file_name,
    "_L_tsvd.RData"
  )
)
rmse_L_tsvdmssa_res[order(rowMeans(rmse_L_tsvdmssa_res[, -1]), decreasing = FALSE), ] |>
  head() |>
  print()
```

## Best L for HO-MSSA
```{r}
R <- 200

plan(multisession)
# plan(sequential)

rmse_L_homssa <- function(R) {
  p <- progressor(R)
  future_replicate(R, {
    p()
    findL_homssa()
  }, simplify = "array", future.seed = 5) |> apply(1:2, mean) |> sqrt()
}

rmse_L_homssa_res <- t(rbind(Ls_ho, rmse_L_homssa(R)))
save(
  rmse_L_homssa_res,
  file = paste0(
    "./Research/comp_results/homssa_",
    save_file_name,
    "_L_tsvd.RData"
  )
)
rmse_L_homssa_res[order(rowMeans(rmse_L_homssa_res[, -1]), decreasing = FALSE), ] |>
  head() |>
  print()
```

## Best L for MSSA
```{r}
R <- 200

plan(multisession)
# plan(sequential)

rmse_L_mssa <- function(R) {
  p <- progressor(R)
  future_replicate(R, {
    p()
    findL_mssa()
  }, simplify = "array", future.seed = 5) |> apply(1:2, mean) |> sqrt()
}

rmse_L_mssa_res <- t(rbind(Ls_mssa, rmse_L_mssa(R)))
save(
  rmse_L_mssa_res,
  file = paste0("./Research/comp_results/mssa_", save_file_name, "_L_tsvd.RData")
)
rmse_L_mssa_res[order(rowMeans(rmse_L_mssa_res[, -1]), decreasing = FALSE), ] |>
  head() |>
  print()
```

## Comparing MV methods
```{r}
R <- 1000

L_tsvd <- rmse_L_tsvdmssa_res[which.min(rowMeans(rmse_L_tsvdmssa_res[, -1])), 1]
L_ho <- rmse_L_homssa_res[which.min(rowMeans(rmse_L_homssa_res[, -1])), 1]
L_mssa <- rmse_L_mssa_res[which.min(rowMeans(rmse_L_mssa_res[, -1])), 1]

all_mv_rmse <- function(R) {
  p <- progressor(R)
  future_replicate(R, {
    p()
    compare_all_mssa()
  }, simplify = "array", future.seed = 5) |> apply(1:2, mean) |> sqrt()
}

all_mv_rmse_res <- all_mv_rmse(R)
save(
  L_mssa,
  L_tsvd,
  L_ho,
  groups_mssa,
  groups_tsvd,
  groups3,
  all_mv_rmse_res,
  sd,
  file = paste0(
    "./Research/comp_results/all_",
    save_file_name,
    "_mv_comp_tsvd.RData"
  )
)
print(all_mv_rmse_res)
```