### 2025-05-14 Code/R/Research/HTLSDstack.Rmd
htlsd methods compare

---
title: "HTLSDstack"
output: html_document
---

```{r setup, include=FALSE}
library(svd)
library(Rssa)
library(rTensor)
library(ggplot2)
library(dplyr)
library(tidyr)
library(future.apply)
library(progressr)

setwd(rprojroot::find_rstudio_root_file())
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
source(file = "./Research/TSSA.R")
```

```{r enable progressr, eval = FALSE}
# This commands should be executed in R console, not as a chunk
# Otherwise they fail
progressr::handlers(global = TRUE)
progressr::handlers("progress")
```

```{r}
rrmse <- function(true, pred) {
  c(sqrt(mean((pred[1, ] - true[1])^2)) * 100 / (0.5 * abs(true[1] - true[2])),
    sqrt(mean((pred[2, ] - true[2])^2)) * 100 / (0.5 * abs(true[1] - true[2])))
}

sd_to_snr <- function(s, sd) {
  10 * log10(sum(abs(s)^2) / sd^2)
}

snr_to_sd <- function(s, snr) {
  sqrt(sum(abs(s)^2) / 10^(snr / 10))
}
```

```{r}
compare_htls <- function() {
  noise <- CCSWGN(N, 0, sd)
  x <- s + noise

  est <- list(
    esprit = parestimate(ssa(x, L = L_esprit), groups = groups_esprit, drop = FALSE),
    htlsd = HTLSDstack(
      x,
      D = D,
      L = L_htlsd,
      groups = groups_htlsd,
      method = "base"
    ),
    tens = HTLSDstack(
      x,
      D = D,
      L = L_tens,
      groups = groups_tens,
      method = "hooi",
      r3 = r3,
      est_dim = est_dim,
      status = FALSE
    )
  )

  sapply(est, function(m)
    sapply(m, function(e)
      c(
        rates = e$rates, frequencies = e$freq
      )))
}
```

# Example from [paper](https://doi.org/10.1002/cem.1212)
```{r}

# Example 1
A <- c(1, 1)
phi <- c(0, 0)
rate <- c(0, 0)
freq <- c(0.2, 0.205)
L_esprit <- 500
L_htlsd <- 50
L_tens <- 50
D <- 10
groups_esprit <- list(1:2)
groups_htlsd <- list(1:2)
groups_tens <- list(1:2)
r3 <- 2
est_dim <- 1
dt <- 0.1
N <- 1000

s <- colSums(A * exp(1i * phi) * exp((-rate + 1i * 2 * pi * freq) %o% seq(
  from = dt,
  length.out = N,
  by = dt
)))
```

```{r warning=FALSE}
sd <- snr_to_sd(s, snr = 30)
R <- 100
with(plan(multisession), {
  library(svd)
batch_estims <- function(R) {
  p <- progressor(R)
  future_replicate(R, {
    p()
    compare_htls()
  }, simplify = "array", future.seed = 5)
}
estims <- batch_estims(R)
})

freq_estims <- apply(estims[3:4,,], 2:3, function(v) c(min(abs(v)), max(abs(v))) / dt)

esprit_rrmse <- apply(freq_estims, 2, rrmse, true = freq)
```