### 2025-09-21 Code/R/Research/cdam25_examples.Rmd
CDAM comments fixes 2

---
title: "CDAM'25 examples"
author: "Khromov Nikita"
output: html_document
---

```{r, setup, include=FALSE}
library(Rssa)
library(scales)
library(rTensor)
library(ggplot2)
library(dplyr)

setwd(rprojroot::find_rstudio_root_file())
knitr::opts_knit$set(root.dir = getwd())
source(file = "./Source/TSSA.R")
source(file = "./Source/cdam25_extra.R")
```

# Single-channel parameters estimation

Loading data
```{r}
load(file = paste0("./Research/comp_results/dim_rrmse_no_rates_tens.RData"))
load(file = paste0("./Research/comp_results/dim_rrmse_no_rates_ssa.RData"))
```

Preparation
```{r}
sd <- 0.04
n <- 25
rates.true <- c(0, 0)
freqs.true <- c(0.2, 0.22)

signal <- exp((rates.true[1] + 2i * pi * freqs.true[1]) * 0:(n - 1)) +
  exp((rates.true[2] + 2i * pi * freqs.true[2]) * 0:(n - 1))

groups <- list(1:2)
max.r <- max(sapply(groups, max))
est.dims <- list(1, 2, 3)
Ls.ssa <- seq(from = max.r + 1, to = (n + 1) %/% 2, by = 1)

# I >= L >= J
Is <- seq(from = max.r + 1, to = n - max.r - 1, by = 1)
Ls <- seq(from = max.r + 1, to = n - max.r - 1, by = 1)

correct.dims <- which(outer(Is, Ls, (
  \(i, l) ifelse((i >= l) &
                   (n - i - l + 2 > max.r) &
                   (n - i - l + 2 <= l), TRUE, FALSE)
)), arr.ind = TRUE)

len.out <- nrow(correct.dims)

tens.dims <- matrix(ncol = 3, nrow = len.out)
colnames(tens.dims) <- c("I", "L", "J")
tens.dims[, "I"] <- Is[correct.dims[,1]]
tens.dims[, "L"] <- Ls[correct.dims[,2]]
tens.dims[, "J"] <- n - tens.dims[, "I"] - tens.dims[, "L"] + 2

tens.dims <- tens.dims[order(tens.dims[,1], tens.dims[,2], decreasing = TRUE),]
```

Table
```{r}
freq1.tssa.df <- freq1.rrmse |> as.data.frame() |>
  mutate(dims = factor(
    x = apply(tens.dims, 1, paste, collapse = "x"),
    levels = apply(tens.dims, 1, paste, collapse = "x")
  )) |>
  pivot_longer(
    !dims,
    names_to = "Estimation\nDimension",
    names_prefix = "V",
    values_to = "RRMSE"
  ) |> mutate(RMSE = RRMSE * freqs.true[1] / 100, Method = "T-SSA")

freq1.ssa.df <- data.frame(RRMSE = freq1.ssa.rrmse, RMSE = freq1.ssa.rrmse * freqs.true[1] / 100, 
                           L = Ls.ssa, Method = rep("SSA", 11))

freq1.summary <- summarise(
  freq1.tssa.df,
  Method = "T-SSA",
  Min = min(RMSE),
  Max = max(RMSE),
  Median = median(RMSE), 
  Mean = mean(RMSE)
) |>
  rbind(
    summarise(
      freq1.ssa.df,
      Method = "SSA",
      Min = min(RMSE),
      Max = max(RMSE),
      Median = median(RMSE),
      Mean = mean(RMSE)
    )
  )

save(freq1.summary, file = "./Research/comp_results/cdam25_1d_param.RData")
xtable(
  freq1.summary,
  digits = 4,
  display = c("s", "s", rep("f", 4)),
  align = rep("c", 6)
)
```

If symmetry relative to L / 2 is considered
```{r}
freq1.summary <- summarise(
  freq1.tssa.df,
  Method = "T-SSA",
  Min = min(RMSE),
  Max = max(RMSE),
  Median = median(RMSE), 
  Mean = mean(RMSE)
) |>
  rbind(
    summarise(
      freq1.ssa.df[7:11, ],
      Method = "SSA",
      Min = min(RMSE),
      Max = max(RMSE),
      Median = median(RMSE),
      Mean = mean(RMSE)
    )
  )

save(freq1.summary, file = "./Research/comp_results/cdam25_1d_param.RData")
xtable(
  freq1.summary,
  digits = 4,
  display = c("s", "s", rep("f", 4)),
  align = rep("c", 6)
)
```

<!-- SSA and T-SSA errors on one plot -->
<!-- ```{r} -->

<!-- x1_range <- range(freq1.ssa.df$L) -->
<!-- x2_range <- range(1:111) -->

<!-- freq1.tssa.df.transformed <- freq1.tssa.df %>% -->
<!--   mutate(Dims_scaled = scales::rescale(1:111, to = x1_range, from = x2_range)) -->

<!-- ggplot() + -->
<!--   geom_line(data = freq1.ssa.df, aes(x = L, y = RMSE, color = "SSA"), size = 1) + -->
<!--   geom_line(data = freq1.tssa.df.transformed, aes(x = Dims_scaled, y = RMSE, color = "T-SSA"), size = 1) + -->
<!--   scale_x_continuous( -->
<!--     name = "L, SSA", -->
<!--     labels = function(x) as.integer(round(x)), -->
<!--     sec.axis = sec_axis( -->
<!--       trans = ~ scales::rescale(., to = x2_range, from = x1_range), -->
<!--       name = "Dimensions index, T-SSA", -->
<!--       labels = function(x) as.integer(round(x)) -->
<!--     ) -->
<!--   ) + -->
<!--   scale_color_manual( -->
<!--     name = "Method", -->
<!--     values = c("SSA" = "blue", "T-SSA" = "red") -->
<!--   ) + -->
<!--   labs( -->
<!--     y = "RMSE", -->
<!--   ) + -->
<!--   theme_minimal() + -->
<!--   theme( -->
<!--     legend.position = "bottom" -->
<!--   ) -->

<!-- ``` -->

# Single Channel signal extraction

Loading data
```{r}
load(file = paste0("./Research/comp_results/signal_no_rates_hossa.RData"))
load(file = paste0("./Research/comp_results/signal_no_rates_ssa.RData"))
```

Preparation
```{r}
sd <- 0.04
n <- 25
rates.true <- c(0, 0)
freqs.true <- c(0.2, 0.22)

signal <- exp((rates.true[1] + 2i * pi * freqs.true[1]) * 0:(n - 1)) +
  exp((rates.true[2] + 2i * pi * freqs.true[2]) * 0:(n - 1))

groups <- list(1:2)
max.r <- max(sapply(groups, max))
trunc_dims <- list(1, 1:2, 1:3)
Ls.ssa <- seq(from = max.r + 1, to = (n + 1) %/% 2, by = 1)

# I >= L >= J
Is <- seq(from = max.r + 1, to = n - max.r - 1, by = 1)
Ls <- seq(from = max.r + 1, to = n - max.r - 1, by = 1)

correct.dims <- which(outer(Is, Ls, (
  \(i, l) ifelse((i >= l) &
                   (n - i - l + 2 > max.r) &
                   (n - i - l + 2 <= l), TRUE, FALSE)
)), arr.ind = TRUE)

len.out <- nrow(correct.dims)

tens.dims <- matrix(ncol = 3, nrow = len.out)
colnames(tens.dims) <- c("I", "L", "J")
tens.dims[, "I"] <- Is[correct.dims[,1]]
tens.dims[, "L"] <- Ls[correct.dims[,2]]
tens.dims[, "J"] <- n - tens.dims[, "I"] - tens.dims[, "L"] + 2

tens.dims <- tens.dims[order(tens.dims[,1], tens.dims[,2], decreasing = TRUE),]

R <- 500
```

Table
```{r}
rec.tssa.df <- rec.rmse |> as.data.frame() |>
  mutate(dims = factor(x = apply(tens.dims, 1, paste, collapse = "x"),
                       levels = apply(tens.dims, 1, paste, collapse = "x"))) |>
  pivot_longer(!dims, names_to = "trunc dims", names_prefix = "V", values_to = "RMSE") |>
  mutate(`trunc dims` = as.factor(paste(trunc_dims[as.integer(`trunc dims`)])))

rec.ssa.df <- data.frame(RMSE = rec.ssa.rmse, L = Ls.ssa)

rec.summary <- summarise(
  rec.tssa.df,
  Method = "T-SSA",
  Min = min(RMSE),
  Max = max(RMSE),
  Median = median(RMSE), 
  Mean = mean(RMSE)
) |>
  rbind(
    summarise(
      rec.ssa.df,
      Method = "SSA",
      Min = min(RMSE),
      Max = max(RMSE),
      Median = median(RMSE),
      Mean = mean(RMSE)
    )
  )

save(rec.summary, file = "./Research/comp_results/cdam25_1d_rec.RData")
xtable(
  rec.summary,
  digits = 4,
  display = c("s", "s", rep("f", 4)),
  align = rep("c", 6)
)
```