### 2025-11-07 Code/R/Research/fast_svd.Rmd
Implement fast hosvd and hooi via primme

---
title: "Testing fast hosvd variants"
output: html_document
---

```{r}
setwd(rprojroot::find_rstudio_root_file())
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

library(Rssa)
library(rTensor)
library(PRIMME)
library(Rcpp)
library(RcppArmadillo)

source(file = "./Source/TSSA.R")
```

# MV case

Original series
```{r}
s <- matrix(1:18 + 1i * 1:18, ncol = 3)
print(s)
```

Construcucting tensor
```{r}
s.tens <- tens3(s, 3, "MSSA")
print(s.tens@data)
```

Computing HOSVD using default SVD
```{r}
s.hosvd.svd <- hosvd_mod(s.tens, svd.method = "svd")
print(s.hosvd.svd$est@data)
```

Computing HOSVD using faster SVD
```{r}
s.hosvd.primme <- hosvd_mod(s.tens, svd.method = "primme")
all.equal(s.hosvd.svd$est@data, s.hosvd.primme$est@data)
```

Computing truncated HOSVD using default SVD
```{r}
s.hosvd.svd.trunc <- hosvd_mod(s.tens, ranks = c(2, 2, 2), svd.method = "svd")
print(s.hosvd.svd.trunc$est@data)
```

Computing truncated HOSVD using faster SVD
```{r}
s.hosvd.primme.trunc <- hosvd_mod(s.tens, ranks = c(2, 2, 2), svd.method = "primme")
all.equal(s.hosvd.svd.trunc$est@data, s.hosvd.primme.trunc$est@data)
```

Computing HOOI-HOSVD using default SVD
```{r}
s.hooi.svd <- tucker_mod(s.tens, ranks = c(3, 4, 3), svd.method = "svd")
print(s.hooi.svd$est@data)
```

Computing HOOI using faster SVD
```{r}
s.hooi.primme <- hosvd_mod(s.tens, ranks = c(3, 4, 3), svd.method = "primme")
all.equal(s.hooi.svd$est@data, s.hooi.primme$est@data)
```

Computing truncated HOOI using default SVD
```{r}
s.hooi.svd.trunc <- tucker_mod(s.tens, ranks = c(2, 2, 2), svd.method = "svd")
print(s.hooi.svd.trunc$est@data)
```

Computing truncated HOSVD using faster SVD
```{r}
s.hooi.primme.trunc <- tucker_mod(s.tens, ranks = c(2, 2, 2), svd.method = "primme")
all.equal(s.hooi.svd.trunc$est@data, s.hooi.primme.trunc$est@data)
```

# 1D case

Original series
```{r}
s.1d <- 1:10 + 1i * 1:10
s.1d.tens <- tens3(s.1d, c(3, 4), "SSA")
print(s.1d.tens@data)
```

Computing truncated HOSVD with default SVD
```{r}
s.1d.hosvd.svd <- hosvd_mod(s.1d.tens, ranks = c(2, 2, 2), svd.method = "svd")
print(s.1d.hosvd.svd$est@data)
```

Computing truncated HOSVD with faster SVD
```{r}
s.1d.hosvd.primme <- hosvd_mod(s.1d.tens, ranks = c(2, 2, 2), svd.method = "primme")
all.equal(s.1d.hosvd.svd$est@data, s.1d.hosvd.primme$est@data)
```

Computing truncated HOOI with default SVD
```{r}
s.1d.hooi.svd <- tucker_mod(s.1d.tens, ranks = c(2, 2, 2), svd.method = "svd")
print(s.1d.hooi.svd$est@data)
```

Computing truncated HOOI with faster SVD
```{r}
s.1d.hooi.primme <- tucker_mod(s.1d.tens, ranks = c(2, 2, 2), svd.method = "primme")
all.equal(s.1d.hooi.svd$est@data, s.1d.hooi.primme$est@data)
```

# Time Benchmarks

```{r}
library(microbenchmark)
```

## MV Case

Short series, HOSVD
```{r}
microbenchmark(hosvd_mod(s.tens, ranks = c(2, 2, 2), svd.method = "svd", status = FALSE), 
               hosvd_mod(s.tens, ranks = c(2, 2, 2), svd.method = "primme", status = FALSE))
```

Short series, HOOI
```{r}
microbenchmark(tucker_mod(s.tens, ranks = c(2, 2, 2), svd.method = "svd", status = FALSE), 
               tucker_mod(s.tens, ranks = c(2, 2, 2), svd.method = "primme", status = FALSE))
```

Long series, HOSVD
```{r}
s.long <- matrix(1:7500 + 1i * 1:7500, ncol = 3)
s.long.tens <- tens3(s.long, 1250, "MSSA")
```

```{r}
cat(
  "Default SVD:\n",
  system.time({hosvd_mod(s.long.tens, ranks = c(2, 2, 2), svd.method = "svd")}),
  "\n\nFaster SVD:\n",
  system.time({hosvd_mod(s.long.tens, ranks = c(2, 2, 2), svd.method = "primme")}))
```

Long series, HOOI
```{r}
cat(
  "Default SVD:\n",
  system.time({tucker_mod(s.long.tens, ranks = c(2, 2, 2), svd.method = "svd")}),
  "\n\nFaster SVD:\n",
  system.time({tucker_mod(s.long.tens, ranks = c(2, 2, 2), svd.method = "primme")}))
```

## 1D Case

Short series, HOSVD
```{r}
microbenchmark(hosvd_mod(s.1d.tens, ranks = c(2, 2, 2), svd.method = "svd", status = FALSE), 
               hosvd_mod(s.1d.tens, ranks = c(2, 2, 2), svd.method = "primme", status = FALSE))
```

Short series, HOOI
```{r}
microbenchmark(tucker_mod(s.1d.tens, ranks = c(2, 2, 2), svd.method = "svd", status = FALSE), 
               tucker_mod(s.1d.tens, ranks = c(2, 2, 2), svd.method = "primme", status = FALSE))
```

Long series, HOSVD
```{r}
s.1d.long <- 1:3001 + 1i * 1:3001
s.1d.long.tens <- tens3(s.1d.long, c(1500, 1500), "SSA")
```

```{r}
cat(
  "Default SVD:\n",
  system.time({hosvd_mod(s.1d.long.tens, ranks = c(2, 2, 2), svd.method = "svd")}),
  "\n\nFaster SVD:\n",
  system.time({hosvd_mod(s.1d.long.tens, ranks = c(2, 2, 2), svd.method = "primme")}))
```

Long series, HOOI
```{r}
cat(
  "Default SVD:\n",
  system.time({tucker_mod(s.1d.long.tens, ranks = c(2, 2, 2), svd.method = "svd")}),
  "\n\nFaster SVD:\n",
  system.time({tucker_mod(s.1d.long.tens, ranks = c(2, 2, 2), svd.method = "primme")}))
```