### 2025-12-29 Code/R/Research/Eigenvalues.Rmd
Add section on R3 research

---
title: "Eigenvalues"
output: html_document
---

```{r, setup, include=FALSE}
library(Rssa)
library(rTensor)
library(ggplot2)
library(english)
library(gridExtra)
library(factoextra)
library(dplyr)
library(tidyr)
root.dir <- rprojroot::find_rstudio_root_file()
knitr::opts_knit$set(root.dir = root.dir)
knitr::opts_chunk$set(root.dir = root.dir)

source(file = sprintf("%s/Source/TSSA.R", root.dir))
source(file = sprintf("%s/Source/tsgen.R", root.dir))
source(file = sprintf("%s/Source/utils.R", root.dir))
```

### Formula for eigenvalues of constant mv series
```{r}
N <- 20
P <- 4
L <- 10
A <- c(1, 1, 1, 3)
s_const <- matrix(rep(A, N * P / length(A)), ncol = N) |> t()
tens3(s_const, L = L, kind = "MSSA") |>
  hosvd_mod(status = FALSE) |>
  get_tensor_eigenvalues() |>
  lapply(zapsmall) |>
  print()
print(sqrt(mean(A^2)) * sqrt(P * L * (N - L + 1)))
```

# Noise eigenvalues
```{r}
set.seed(5)
# N <- 19; L <- 7
N <- 967; L <- 443
# N <- 967; L <- 484
# N <- 19; L <- 10
# N <- 99; L <- 20
P <- 200
K <- N - L + 1
modes <- c(L, K, P)
noise <- matrix(rnorm(N * P), ncol = P)
noise_tens <- tens3(noise, L, "mssa")
noise_tens_fnorm <- fnorm_complex(noise_tens)
sd_est <- sqrt(mean(noise^2))
sd_tens_est <- sqrt(mean(noise_tens@data^2))
ev <- noise |> tens_mssa_decompose(L, neig = modes, status = FALSE) |>
      get_tensor_eigenvalues()
ev_sums <- ev |> sapply(\(x) sum(x^2)) |> sqrt()

ssa_ev <- ssa(noise, L, kind = "mssa", svd.method = "primme", neig = min(L, K * P))$sigma
ssa_ev_sum <- sqrt(sum(ssa_ev^2))

cat("Tensor eigenvalues:\n")
invisible(lapply(seq_along(ev), \(i) {
  cat("  Dim", i, "-", ev[[i]], '\n')
  cat(
    "  Marchenko–Pastur Bounds for eigenvalues:",
    noise_tens_fnorm * (1 / sqrt(min(modes[i], prod(modes[-i]))) + 1 / sqrt(max(modes[i], prod(modes[-i]))) * c(1, -1)),
    "(actual:", sprintf("%f, %f)", max(ev[[i]]), min(ev[[i]])),
    "\n"
  )
}))
cat("Tensor eigenvalue sums:", ev_sums, '\n')
cat("Matrix eigenvalues:\n ", ssa_ev, '\n')
cat("Matrix eigenvalue sum:", ssa_ev_sum, '\n')
cat("Tensor/Matrix F norm:", noise_tens_fnorm, '\n')
cat("RMS for ts:", sd_est, '\n')
cat("RMS for tensor:", sd_tens_est, '\n')
cat("TEVS / RMS =", ev_sums / sd_est, '\n')
cat("TEVS / RMS_tens =", ev_sums / sd_tens_est, '\n')
```

# Noise mixing for MSSA and HO-MSSA

```{r}
set.seed(5)
sds <- seq(from = 0, to = 1.2, by = 0.2)

# S.true <- 1
S.true <- 2
S_case <- sprintf("%s_periodic%s", as.english(S.true), ifelse(S.true > 1, "s", ""))

# P <- 2
# P <- 4
# P <- 8
# P <- 12
# P <- 16
# P <- 20
P <- 40
P_case <- sprintf("_c%d", P)

# rates.true <- matrix(rep(0, P * S.true), nrow = P)
# rates_case <- "_nr"
# rates.true <- matrix(rep(-0.04, P * S.true), nrow = P)
# rates_case <- "_Ler"
rates.true <- matrix(rep(-0.04, P * S.true) + rep(seq(
  from = 0,
  length.out = S.true,
  by = -0.02
), each = P), nrow = P)
rates_case <- "_dsr"

# freqs.true <- matrix(rep(0.2, P * S.true), nrow = P)
# freqs_case <- "_ef"
freqs.true <- rep(1/8, P) %o% seq(from = 1, to = 8 / 6, length.out = S.true)
freqs_case <- "_dsf"

# A <- matrix(rnorm(P * S.true), nrow = P)
# A_case <- "_da"
# A <- matrix(seq(-2, 2, length.out = P * S.true), nrow = P)
# A_case <- "_ua"
A <- matrix(rep(1, P * S.true), nrow = P)
A_case <- "_ea"

poly.true <- 1 %o% rep(1, P) %o% rep(1, S.true)
poly_case <- ""
# poly.true <- c(1, 1) * matrix(runif(2 * P, 0.5, 2), nrow = 2) %o% rnorm(S.true, mean = 1, sd = 0.25)
# poly.true <- c(1, 1, 1) * matrix(runif(3 * P, 0.5, 2), nrow = 3) %o% rnorm(S.true, mean = 1, sd = 0.25)
# poly.true <- c(1, 1) * matrix(1:(2*P), nrow = 2) %o% rnorm(S.true, mean = 1, sd = 0.25)
# poly_case <- sprintf("p%d", dim(poly.true)[1] - 1)

# phases.true <- matrix(rep(0, P * S.true), nrow = P)
# phases_case <- "_ep"
phases.true <- rep(1, P) %o% runif(S.true, 0, 2 * pi)
phases_case <- "_dp"

# complex.signal <- TRUE
# complex_case <- "_c" # complex
complex.signal <- FALSE
complex_case <- "_r" # real

# N <- 25
# n_case <- ""
N <- 99
n_case <- "_long"
```

```{r}
full_case <- paste0(n_case, P_case, rates_case, freqs_case, A_case, poly_case, phases_case, complex_case)
sample_step = 100 / N
max.r <- calc_rank(A, rates.true, freqs.true * sample_step, phases.true, complex.signal, poly.true)

stopifnot(N > max.r * 2 + 1)
signal <- construct_ts(N, A, rates.true, freqs.true, phases.true, complex.signal, poly.true, by = sample_step)
max.r3 <- calc_rank3(signal)
```

## Comparing smallest signal eigenvalues to noise eigenvalues

```{r}
Ls_mssa <- seq.int(from = max.r + 1, to = N - max.r)
Ls_homssa <- seq.int(from = max.r + 1, to = (N + 1) %/% 2)
Ks_mssa <- N - Ls_mssa + 1
Ks_homssa <- N - Ls_homssa + 1

sigmas_mssa <- sapply(Ls_mssa, \(L) ssa(signal, L, kind = "mssa")$sigma[max.r])
sigmas_homssa <- sapply(Ls_homssa, \(L) {
  tens_ev <- get_tensor_eigenvalues(tens_mssa_decompose(signal, L, "hooi", status = FALSE))
  c(homssa1 = tens_ev[[1]][max.r],
    homssa2 = tens_ev[[2]][max.r],
    homssa3 = tens_ev[[3]][max.r3])
})

## All eigenvalues of the signal for particular L:
L <- (N + 1) %/% 2
get_tensor_eigenvalues(tens_mssa_decompose(signal, L, "hosvd", status = FALSE))
```

```{r}
mix_sd2_mssa <- sigmas_mssa^2 * (pmin(Ls_mssa, Ks_mssa * P)) / (Ls_mssa * Ks_mssa * P)
mix_sd2_homssa <- sigmas_homssa^2 * rbind(Ls_homssa, Ks_homssa, P) / (rep(1, 3) %o% (Ls_homssa * Ks_homssa * P))
maxind_mssa <- which.max(mix_sd2_mssa)
maxind_homssa <- which.max(colSums(sqrt(mix_sd2_homssa)))
# maxind_homssa <- which.max(colSums(mix_sd2_homssa))
# maxind_homssa <- which.max(mix_sd2_homssa[2,])
# maxind_homssa <- length(Ls_homssa)
# maxind_homssa <- which.max(apply(mix_sd2_homssa, 2, \(v) -sd(v)))

L_mssa <- Ls_mssa[maxind_mssa]; K_mssa <- Ks_mssa[maxind_mssa]
L_homssa <- Ls_homssa[maxind_homssa]; K_homssa <- Ks_homssa[maxind_homssa]

maxmin_mixsd2 <- c(mssa = mix_sd2_mssa[maxind_mssa],
                   mix_sd2_homssa[, maxind_homssa],
                   homssa_mean = mean(sqrt(mix_sd2_homssa[, maxind_homssa]))^2)

sqrt(maxmin_mixsd2)
```

## Using mix rates

```{r}
if (N == 25) {
  L_mssa <- 19; L_homssa <- 9
} else {
  L_mssa <- 80; L_homssa <- 30
}
R <- 100
pb <- txtProgressBar(0, R * length(sds), style = 3)
last_signal_evs <- sapply(sds, \(sd) {
  set.seed(5)
  apply(replicate(R, {
    setTxtProgressBar(pb, pb$getVal() + 1)
    x <- signal + rnorm(N * P, sd = sd)
    c(mssa = mixing_rate(ssa(x, L_mssa, kind = "mssa"), max.r),
      homssa = mixing_rate(get_tensor_eigenvalues(tens_mssa_decompose(x, L_homssa, "hooi", status = FALSE)), c(max.r, max.r, max.r3)))
  }), 1, mean)
})
close(pb)
```

```{r}
mr_df <- mr |> t() |> data.frame() |> mutate(sd = sds)
mr_long <- mr_df |> pivot_longer(c("mssa", "homssa1", "homssa2", "homssa3"))

mr_plot <- mr_long |> ggplot() + geom_line(aes(x = sd, y = value, color = name), na.rm = TRUE) + ggtitle(sprintf("P = %d", P)) + theme_minimal()
mr_plot_wo3 <- mr_long |> filter(name != "homssa3") |> ggplot() + geom_line(aes(x = sd, y = value, color = name), na.rm = TRUE) + ggtitle(sprintf("P = %d", P)) + theme_minimal()
{
  pdf(
    paste0(
      "./Research/comp_results/cases_errors/noise_mix_plots/mr",
      full_case,
      ".pdf"
    ),
    width = 8,
    height = 6
  )
  plot(mr_plot)
  dev.off()
  pdf(
    paste0(
      "./Research/comp_results/cases_errors/noise_mix_plots/mr_wo3",
      full_case,
      ".pdf"
    ),
    width = 8,
    height = 6
  )
  plot(mr_plot_wo3)
  dev.off()
}
mr_plot
mr_plot_wo3
```